\LARGE \bf Modeling and Control of a Flexible Catheter for Cardiac Ablation

### 2018-05-01 manuscript/manuscript.tex
revisions to mpc analysis. Beginning to work on report documentation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{\LARGE \bf Eliminating Nonzero Initial States in \\Flexible Systems through Specified Insensitivity Input Shaping}

\author{Daniel Newman$^{1}$, Seong-Wook Hong$^{2}$, and Joshua Vaughan$^{1}$% <-this % stops a space
\thanks{$^{1}$Department of Mechanical Engineering,
        University of Louisiana at Lafayette, Lafayette, Louisiana 70504 USA
        {\tt\small joshua.vaughan@louisiana.edu}}%
\thanks{$^{2}$Kumoh National Institute of Technology,
        Gumi, Gyeongbuk 730-701 Republic of Korea}%
}

%\author{Daniel Newman$^{1}$, Joshua Vaughan$^{1}$, and Seong-Wook Hong$^{2}$% <-this % stops a space
%\thanks{*This work was not supported by any organization}% <-this % stops a space
%\thanks{$^{1}$Department of Mechanical Engineering, University of Louisiana at Lafayette, Lafayette, Louisiana 70504, United States of America
%    {\tt\small joshua.vaughan@louisiana.edu},
%        $^{2}$Department of Mechatronics, Kumoh National Institute of Technology, Gumi, Gyeongbuk 730-701, Republic of Korea}
%}%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamic Model}
\label{sec:intro}
%
\begin{figure}[t!] 
\begin{center}
\includegraphics[width = 2.7in]{Figures/Planar_Model} 
\caption{Dynamic model of a two link flexible-arm manipulator} 
\label{fig:model}
\end{center}
\vspace{0.1in}
\end{figure}
%
The dynamic model used in this analysis is shown in Fig.~\ref{fig:model}. A flexible beam is assumed to be fixed at its base. The rotation of each element about the $\bar{Z}$-axis is designated by $\theta_{i}$, where $i$ is the element number. As depicted in this Figure, each rotation is defined relative to the global coordinate frame. The generalized coordinates are arranged in the following form:
%
\begin{equation}
\label{eq:nodaldisp}
Q_{(n \times 1} = 
\begin{bmatrix}
 \theta_{1} & 
 \theta_{2} &
 \dots & 
 \theta_{n} & 
\end{bmatrix}^T.
\end{equation}
%

@@

The input to the system is given as force applied to the distal tip of the $n^{th}$ element, $f$. In addition, a known velocity field, $u$ exerts drag forces on each element. This model serves as a first approximation of hemodynamic forces, where the drag force, $F_D$, exerted on a catheter element is given by:
%
\begin{equation}
\label{eq:drag_force}
F_D = \int_{0}^{l_i} \frac{1}{2}\rho u  C_d D \mathrm{d}l,
\end{equation}
%
where $\rho$ is the mass density of blood, $C_d$ is the drag coefficient of the catheter, $D$ is the diameter of its cross-section, and $l$ is the projected length of the catheter element along the $\bar{X}$-axis. Note that this model ignores wall shear. 
Force $F$ has components $f_x$, $f_y$, and $f_z$ defined in the inertial coordinate frame and yields resultant torques on the beam elements. This resultant torque is defined as:
%
\begin{equation}
\label{eq:forced_torque}
\vec{\gamma}_{n} = \int_{0}^{l}\vec{r} \times \vec{F} \mathrm{d}l,
\end{equation}
%
where $\vec{\gamma}_n$ gives the vector of torques applied about each axis within the $n$ cooordinate frame. The integral denotes the summation of each torque on the continuous length of the beam element as a function of its length, $l$. The vector $\vec{r}$ points along the beam within its local coordinate frame. Because each beam element is aligned with this frame, $\vec{r}$ takes the form:
%
\begin{equation}
\label{eq:r_vector}
\vec{r} = l \mathbf{\hat{i}} + 0 \mathbf{\hat{j}} + \mathbf{\hat{k}},
\end{equation}
%
the resolved force can only exert a moment about the $\mathbf{\hat{j}}$ and $\mathbf{\hat{k}}$ axes in this model. The resulting total exogenous input is given by:
%
\begin{equation}
\label{eq:Gamma}
\Gamma_{(3*n - 2) \times 1} = 
\begin{bmatrix}
 0 & 
 0 &
 \gamma_{y,2} & 
 \gamma_{z,2} &  
 \dots & 
 0 & 
 \gamma_{y,n} & 
 \gamma_{z,n}
\end{bmatrix}^T,
\end{equation}
%
where the shape is once again dictated by the nature of the generalized coordinates, (\ref{eq:nodaldisp}).

@@

Aside from the rotation of the first element, $\theta_{x,1}$, each generalized coordinate has an associated internal moment based on its material properties. An axial torque resists rotation about the $x_n$-axis by the following relation:
%
\begin{equation}
\label{eq:m_x}
M_{x,n} = \frac{GJ (\theta_{x,n} - \theta_{x,n-1})}{l_n + l_{n-1}}.
\end{equation}
%

Additionally, each element has an internal moment related to bending, given by:
%
\begin{equation}
\begin{aligned}
\label{eq:m_y}
M_{y,n} &= \frac{EI (\theta_{y,n} - \theta_{y,n-1})}{l_n + l_{n-1}}, \\
\label{eq:m_z}
M_{z,n} &= \frac{EI (\theta_{z,n} - \theta_{z,n-1})}{l_n + l_{n-1}}. \\
\end{aligned}
\end{equation}
%
These moments are found by discretizing standard Euler-Bernoulli beam theory.

@@

The resulting equations of motion were created by Kane's method using the Python SymPy module. The state-space representation of the dynamic system is given by:
%
\begin{equation}
\label{eq:eqmotion}
\begin{bmatrix}
 \dot{Q}\\
 \ddot{Q}
\end{bmatrix}
=
\begin{bmatrix}
 0 & I \\
 M^{-1} K & 0
\end{bmatrix} 
\begin{bmatrix}
 Q\\
 \dot{Q}
\end{bmatrix}
+
\begin{bmatrix}
0 \\
\Gamma
\end{bmatrix}
 + B u,
\end{equation} 
%
where,
%
\begin{equation}
\label{eq:B_matrix}
B_{2(3n-2) \times 3} = 
\begin{bmatrix}
0_{3n - 2 \times 1} & 1 & 0_{(3n - 1) \times 1} \\
0_{6n - 2\times 1} & 0 & 1
\end{bmatrix}^T,
\end{equation}
%
where the nonzero elements of $B$ correspond to the control inputs $\tau_1$ and $\tau_2$, respectively.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{\LARGE \bf Modeling and Control of a Flexible Catheter for Cardiac Ablation}

\author{Daniel Newman$^{1}$, Charles Taylor$^{1}$, and Joshua Vaughan$^{1}$% <-this % stops a space
\thanks{$^{1}$Department of Mechanical Engineering,
        University of Louisiana at Lafayette, Lafayette, Louisiana 70504 USA
        {\tt\small joshua.vaughan@louisiana.edu}}%
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamic Model}
\label{sec:intro}
%
\begin{figure}[t!] 
\begin{center}
\includegraphics[width = 2.7in]{Figures/Planar_Model} 
\caption{Dynamic model of a two link flexible-arm manipulator} 
\label{fig:model}
\end{center}
\vspace{0.1in}
\end{figure}
%
The dynamic model used in this analysis is shown in Fig.~\ref{fig:model}. A flexible beam is assumed to be fixed at its base. The rotation of each element about the $\bar{Z}$-axis is designated by $\theta_{i}$, where $i$ is the element number. As depicted in this Figure, each rotation is defined relative to the global coordinate frame. The generalized coordinates are arranged in the following form:
%
\begin{equation}
\label{eq:nodaldisp}
Q_{n \times 1} = 
\begin{bmatrix}
 \theta_{1} & 
 \theta_{2} &
 \dots & 
 \theta_{n} & 
\end{bmatrix}^T.
\end{equation}
%

@@

The input to the system is given as force applied to the distal tip of the $n^{th}$ element, $f$. In addition, a known velocity field, $\vec{U}$, exerts drag forces on each element. These drag forces generate moments at each hinge, causing deflection of the catheter. These moments are given by:
%
\begin{equation}
\vec{\gamma}_{n} = \int_{0}^{l_n} \vec{r}_n(x) \times \vec{f}_n(x) \mathrm{d}x,
\end{equation}
%
where:
%
\begin{equation}
\label{eq:r_vector}
\vec{r}(x)_n = x \mathbf{\hat{i}_n} + 0 \mathbf{\hat{j}_n} + \mathbf{\hat{k}_n},
\end{equation}
%
is a displacement vector pointing from the element origin to a point along its length, $\vec{f}_n(x)$ is the location-dependent force vector due to hemodynamic forces, and $\vec{\gamma}_{n}$, is the resultant torque acting at the element origin.

The drag induced by hemodynamic flow is:
%
\begin{equation}
\vec{f}_n(x,t) = \frac{1}{2}\rho C_d D 
\begin{bmatrix}
\cos{\theta_n} & -\sin{\theta_n} \\
\sin{\theta_n} & \cos{\theta_n}
\end{bmatrix}
\vec{U}(x,t),
\end{equation}
where $\rho$ is the mass density of blood, $C_d$ is the drag coefficient of the catheter, $D$ is the diameter of its cross-section, and $\vec{U}(x)$ is the flow field, given in the inertial coordinate frame. Note that this model ignores wall shear. The resulting total exogenous input is therefore given by:
%
\begin{equation}
\label{eq:Gamma}
\Gamma_{n \times 1} = 
\begin{bmatrix}
\gamma_{2} &  
\dots & 
\gamma_{n-1} & 
\gamma_{n}
\end{bmatrix}^T,
\end{equation}
%
where the torque applied to the first element is ignored due to the fixed boundary condition.

@@

Each generalized coordinate also has an associated internal moment based on its material properties. The internal bending moment for each element is given by:
%
\begin{equation}
\begin{aligned}
\label{eq:m_z}
M_{n} &= \frac{EI (\theta_{n} - \theta_{n-1})}{l_n + l_{n-1}}, \\
\end{aligned}
\end{equation}
%
These moments are found by discretizing standard Euler-Bernoulli beam theory.

@@

The resulting equations of motion were created by Kane's method using the Python SymPy module. When linearized about $Q=\dot{Q}=0$, the state-space representation of the dynamic system is given by:
%
\begin{equation}
\label{eq:eqmotion}
\begin{bmatrix}
 \dot{Q}\\
 \ddot{Q}
\end{bmatrix}
=
A
\begin{bmatrix}
 Q\\
 \dot{Q}
\end{bmatrix}
+
B f_{in}
+
\Omega \Gamma

\end{equation} 
%
where $A$ and $B$ are the linearized state transition and input matrices, and $\Omega$ is the disturbance input matrix.

Because (\ref{eq:eqmotion}) is linearized, it will degrade in performance at large tip deflections. This degradation is increased due to the use of Euler-Bernoulli beam theory, which is strictly valid for small deflections. However, this model serves as a foundation from which dynamic and control properties can be evaluated.

\section{Control Method}
\label{sec:control}

The control objective for this initial study is to hold the catheter endpoint steady at a specified point in $\bar{X}$. Due the the external disturbance induced by blood flow, a model predictive control (MPC) approach will be used to maintain accurate positioning. The general formulation for MPC of a linear system is realized by discretizing the system:
%
\begin{equation}
\begin{align}
\label{eq:discrete-ss}
x_{n+k} &= A x_n + B u_k + \Omega \Gamma_k\\
y_n &= C x_k,
\end{align}
\end{equation}
%
and determining the correct input, $u_k$ which minimizes a quadratic cost function:
%
\begin{equation}
